Learning words by recursive pragmatic reasoning about other agents

### 2013-05-31 praglearn.tex
text

The acquisition of scalar adjectives like ``some'' provides a puzzle for most models of word learning: given that in many contexts, the word ``some'' is used to mean \textsc{some but not all}, how do children learn that this is not in fact its literal meaning? Our model is able to take specificity implicatures into account when learning, and thus potentially answer this puzzle. Imagine two bowls: one containing a mixture of apples and oranges, the other containing apples alone. Fig.~\ref{fig:scalar} (left) shows the model's beliefs after encountering the words ``some'' and ``all'' used contrastively to refer to these two bowls (as in e.g.\ ``the bowl with some apples''). It believes that at least one of these words must have a restrictive literal meaning, but not both -- it leaves open the possibility that either ``some'' or ``all'' might be literally ambiguous. Fig.~\ref{fig:scalar} (right) then shows the model's beliefs if we also provide examples of ``some'' referring to exactly the same object, but in contexts where this is the only object available, so there is no contrast and thus no implicature; this convinces the model that ``all'' literally refers only to the bowl that contains only apples, while ``some'' most likely is somewhat ambiguous.

>>>>

The acquisition of scalar adjectives like ``some'' provides a puzzle for most models of word learning: given that in many contexts, the word ``some'' is used to mean \textsc{some but not all}, how do children learn that this is not in fact its literal meaning? Our model is able to take specificity implicatures into account when learning, and thus potentially answer this puzzle. Imagine two bowls: one containing a mixture of apples and oranges, the other containing apples alone. Fig.~\ref{fig:scalar} (left) shows the model's beliefs after encountering the words ``some'' and ``all'' used contrastively to refer to these two bowls (as in e.g.\ ``the bowl with some apples''). It believes that at least one of these words must have a restrictive literal meaning, but not both -- it leaves open the possibility that either ``some'' or ``all'' might be literally ambiguous. Fig.~\ref{fig:scalar} (right) then shows the model's beliefs if we also provide examples of ``some'' referring to exactly the same object, but in contexts where this is the only object available, so there is no contrast and thus no implicature; this convinces the model that ``all apple'' literally refers only to the bowl that contains no oranges, while ``some'' most likely is somewhat ambiguous. Crucially, the model is able to learn this without once ever observing ``some'' used
